---
title: "viz i"
output: github_document
---

```{r setup}
library(tidyverse)
```

```{r}
weather_df=read_csv("./weather_df.csv", na=c("NA","","."))

weather_df
```

##scatterplots
create my first scatterplot ever
```{r}
ggplot(weather_df,aes(x = tmin, y = tmax)) +
  geom_point()
```

new approach , same plot
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax)) +
  geom_point()
```
save ad edit a plot object
```{r}
weather_plot=
  weather_df %>% 
  ggplot(aes(x= tmin, y =tmax))

weather_plot+geom_point()
```

## make an advance scatterplot
start with the same one and make it fancy
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax, color=name)) +
  geom_point()+
  geom_smooth(se=FALSE)
```

## what about the aes placement
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax)) +
  geom_point(aes(color=name))+
  geom_smooth()
```

noew facet
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax, color=name)) +
  geom_point()+
  geom_smooth(se=FALSE, alpha= 0.5, size=0.2)+
  facet_grid(. ~ name)
```
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax, color=name)) +
  geom_point()+
  geom_smooth(se=FALSE)+
  facet_grid(name ~ .)
```

let's combine some elements and try a new plot
```{r}
weather_df %>% 
  ggplot(aes(x=date, y =tmax, color=name)) +
  geom_point(alpha=0.5,aes(size=prcp))+
  geom_smooth(se=FALSE)+
  facet_grid(. ~ name)
```

## small notes
 how many geoms have to exsist?
 you can have whatever geoms you want. But we need to include the points(better viewment)
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax, color=name)) +
  geom_smooth(se= FALSE)
```

you can use a neat geom
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax)) +
  geom_hex()
```

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax)) +
  geom_bin2d()
```

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y =tmax)) +
  geom_density2d()
```

## univariate plots
histograms are great 
```{r}
weather_df %>% 
  ggplot(aes(x= tmin))+
  geom_histogram()
```


```{r}
weather_df %>% 
  ggplot(aes(x= tmin, fill =name))+
  geom_histogram(position="dodge")+
  facet_grid(.~name)
```

```{r}
ggplot(weather_df, aes(x = tmin, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
## Warning: Removed 17 rows containing non-finite outside the scale range## (`stat_density()`).
```



boxplots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_boxplot()
## Warning: Removed 17 rows containing non-finite outside the scale range## (`stat_boxplot()`).
```

## trendy plots
violin plots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```
```{r}
library(ggridges)
```

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges()
## Picking joint bandwidth of 1.54## Warning: Removed 17 rows containing non-finite outside the scale range## (`stat_density_ridges()`).
```

```{r}
ggp_weather =
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 
ggp_weather 

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```

```{r}
ggp_weather %>% 
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

ggp_weather
```

```{r fig.width = 6, fig.asp = .6, }
ggp_weather
```















